---
title: "Mhorseshoe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mhorseshoe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(Mhorseshoe)
library(ggplot2)
library(horseshoe)
```

In this vignette, the functions exact_horseshoe, approx_horseshoe are explained 
in the following sections.

1. About the horseshoe estimator in this package
2. Brief Description of approximate MCMC algorithm

# 1. About the horseshoe estimator in this package

The horseshoe prior is a continuous reduced prior distribution frequently used 
in high-dimensional Bayesian linear models, and can be effectively applied to 
sparse data. This is a method that theoretically guarantees excellent shrinkage 
properties. Mhorseshoe provides an estimator applying the horseshoe prior, The 
likelihood function of the Gaussian linear model to be considered in this 
package is as follows:

$$L(y\ |\ x, \beta, \sigma^2) = (\frac{1}{\sqrt{2\pi}\sigma})^{-N/2}exp
\{ -\frac{1}{2\sigma^2}(y-X\beta)^T(y-X\beta)\},\\ X \in 
\mathbb{R}^{N \times p},\ y \in \mathbb{R}^{N},\ \beta \in \mathbb{R}^{p}$$

And the form of the hierarchical model applying the horseshoe prior to $\beta$ 
is expressed as follows:

$$\beta_{j}\ |\ \sigma^{2}, \tau^{2}, \lambda_{j}^{2}
\overset{i.i.d}{\sim} N\left(0, \sigma^{2}\tau^{2}\lambda_{j}^{2} \right),
\quad \lambda_{j}\overset{i.i.d}{\sim}C^{+}(0,1),\quad j=1,2,...,p, \\ \tau
\overset{i.i.d}{\sim}C^{+}(0,1),\quad p(\sigma^2)\propto gamma\left(
w/2, w/2\right).$$

Where $C^{+}(0,1)$ is a half-Cauchy distribution, 
$\lambda_{1}, ..., \lambda_{p}$ are local shrinkage parameters, and $\tau$ is 
global shrinkage parameter. The exact_horseshoe in this package is a horseshoe 
estimator for the prior defined above, as described in exact MCMC algorithm of 
Johndrow et al. (2020). The blocked Metropolis-within-Gibb that samples 
$\lambda,\ \left(\tau, \sigma, \beta \right)$ is adopted, and the brief process 
is as follows.

$$\lambda=(\lambda_1, \lambda_2, ..., \lambda_p)^T,\quad \xi=\tau^{-2}, \\ 
\Lambda = diag\left(\lambda_{1}^{2},..., \lambda_{p}^{2} \right),
\quad M = I_{N} + \tau^{2}X\Lambda X^{T}, \\ p(\xi | y, X, \lambda) 
\propto |M |^{-1/2}\{(w + y^{T}M^{-1}y) / 2 \}^{-(w + N) / 2}/
\{\sqrt{\xi}(1+\xi)\}.$$

1. Sample $\lambda_{j}^{-2} = \eta_{j},\ j=1,2,...,p$, using the
following posterior of $\eta_{j}$ for rejection sampling.
$$p(\eta_{j} | \tau, \beta_{j}, \sigma) \propto \frac{1}{1 +
\eta_{j}}exp\{-\frac{\beta_{j}^{2}\xi \eta_{j}}{2\sigma^{2}} \}.$$
2. Sample $\tau^{-2} = \xi$ using the proposed MH algorithm as follows.
$$log(\xi^{\star}) \sim N(log(\xi), s), accept\ \xi \ w.p.\
\{p(\xi^{\star} | y, X, \lambda)\xi^{\star}\}/\{p(\xi | y, X, \lambda)\xi \}.$$
3. Sample $\sigma^{2}$, $$\sigma^{2} | y, X, \lambda, \tau \sim
InvGamma\{(w + N) / 2, (w + y^{T}M^{-1}y) / 2 \}.$$
4. Sample $\beta$, $$\beta | y, X, \lambda, \tau, \sigma \sim
N\left(\left(X^{T}X + \left(\tau^{2} \Lambda \right)^{-1}\right)^{-1}X^{T}y,
\ \sigma^{2}\left(X^{T}X + \left(\tau^{2} \Lambda \right)^{-1} \right)^{-1} \right).$$

Since the algorithm of this package considers the case of $p >> N$, the 
computational cost can be lowered by sampling $\beta$ using fast 
sampling(Bhattacharya et al.,2016) in the step 4.

$$u \sim N_{p}(0, \tau^{2}\Lambda),\quad f \sim N_{N}(0, I_{N}), \\
v = Xu + f,\quad v^{\star} = M^{-1}(y/\sigma - v), \\
\beta = \sigma(u + \tau^2 \Lambda X^{T}v^{\star}).$$

For simulation, the data size was set to N = 300, p = 500, and the design 
matrix $X=\{x_{j} \}_{j=1}^{p}$, true value of beta were set as follows.

$$x_{j} \sim N_{N}\left(0, I_{N} \right), \\ y_{i} \sim N(x_{i}\beta, 4), \\ 
\beta_{j} = \begin{cases} 1, & \mbox{if }\ \mbox{j<51,} \\ 0, & 
\mbox{otherwise.}\end{cases}$$

As a caution, the data is for testing the application of the algorithm and is 
not a simulation that strictly considers sparsity condition.

```{r}
# making simulation data.
set.seed(123)
N <- 300
p <- 500
p_star <- 50
true_beta <- c(rep(1, p_star), rep(0, p-p_star))

# design matrix X.
X <- matrix(1, nrow = N, ncol = p)
for (i in 1:p) {
  X[, i] <- stats::rnorm(N, mean = 0, sd = 1)
}
  
# response variable y.
y <- vector(mode = "numeric", length = N)
e <- rnorm(N, mean = 0, sd = 2)
for (i in 1:p_star) {
  y <- y + true_beta[i] * X[, i]
}
y <- y + e
```

For the corresponding simulation data, the results were compared with the 
horseshoe function of the horseshoe package. We fit the data to the 
horseshoe and exact_horseshoe functions and plot the results for the first 100 
indices, including the first 50 indices that are non-zero coefficients.

```{r fig.width = 6, fig.height= 4}
# horseshoe in horseshoe package.
horseshoe_result <- horseshoe::horseshoe(y, X, method.tau = "halfCauchy",
                                         method.sigma = "Jeffreys",
                                         burn = 0, nmc = 500)

# exact_horseshoe.
exact_horseshoe_result <- exact_horseshoe(y, X, burn = 0, iter = 500)

df <- data.frame(index = 1:100,
                 horseshoe_BetaHat = horseshoe_result$BetaHat[1:100],
                 horseshoe_LeftCI = horseshoe_result$LeftCI[1:100],
                 horseshoe_RightCI = horseshoe_result$RightCI[1:100],
                 exhorseshoe_BetaHat = exact_horseshoe_result$BetaHat[1:100],
                 exhorseshoe_LeftCI = exact_horseshoe_result$LeftCI[1:100],
                 exhorseshoe_RightCI = exact_horseshoe_result$RightCI[1:100],
                 true_beta = true_beta[1:100])

# Estimation results of the horseshoe in horseshoe package.
ggplot(data = df, aes(x = index, y = true_beta)) + 
  geom_point(size = 2) + 
  geom_point(aes(x = index, y = horseshoe_BetaHat), size = 2, col = "red") +
  geom_errorbar(aes(ymin = horseshoe_LeftCI,
                    ymax = horseshoe_RightCI), width = .1, col = "red") +
  labs(title = "95% Credible intervals of the horseshoe in horseshoe package", 
       y = "beta")

# Estimation results of the exact_horseshoe in Mhorseshoe package.
ggplot(data = df, aes(x = index, y = true_beta)) + 
  geom_point(size = 2) + 
  geom_point(aes(x = index, y = exhorseshoe_BetaHat), 
             size = 2, col = "red") +
  geom_errorbar(aes(ymin = exhorseshoe_LeftCI, 
                    ymax = exhorseshoe_RightCI), width = .1, col = "red") +
  labs(title = "95% Credible intervals of the exact_horseshoe in Mhorseshoe 
       package", y = "beta")
```

The results for both functions were similar. Both functions estimate that 95% 
of the credible intervals for all non-zero indices except 4, 36, 41, and 46 include 1.

# 2. approximate MCMC algorithm

Johndrow et al. (2020) proposed a scalable approximate MCMC algorithm that can 
reduce computational costs by introducing a thresholding method while applying 
the horseshoe prior. The set of columns that satisfies the condition 
($\tau^{2}\lambda_{j}^{2} > \delta$) is defined as the active set, and let's 
define $S$ as the set of indices of columns that satisfy the condition.

$$S = \{j\ |\ \tau^{2}\lambda_{j}^{2} > \delta,\ j=1,2,...,p. \}.$$

If $\tau^{2}\lambda_{j}^{2}$ is very small, the posterior of $\beta$
will have a mean and variance close to 0. Let the number of elements 
of $S$ be $s_{\delta}$.

$$X_{S} \in R^{N \times s_{\delta}},
\quad \Lambda_{S} \in R^{s_{\delta} \times s_{\delta}}, \\ M_{\delta}^{-1} =
\left(I_{N} + \xi^{-1}X_{S}\Lambda_{S}X_{S}^{T} \right)^{-1}.$$

$M_{\delta}^{-1}$, which reduces the computational cost, is applied to all 
parts of the algorithm, $\beta$ samples are extracted from the posterior using 
fast sampling(Bhattacharya et al.,2016). The above changed algorithm is 
implemented in the approx_horseshoe function. Johndrow et al. (2020) uses a 
fixed $\delta$ in the approximate algorithm, On the other hand, our package 
applies an algorithm that sets $\delta$ with the adaptive probability.

And we adopt $\delta$ through adaptive probability.

The adaptive algorithm developed in this package estimates a new threshold 
$\delta^{New}$ with updated shrinkage parameters by MCMC sampling, Adopt 
$\delta^{New}$ through adaptive probability. the process of updating a new 
threshold is added by applying the properties of the shrinkage weight 
$k_{j},\ j=1,2,...,p$ proposed by Piironen and Vehtari (2017). In the prior of
$\beta_{j} \sim N(0, \sigma^{2}\tau^{2}\lambda_{j}^{2}) = N(0, \sigma^{2}\xi^{-1}\eta_{j}^{-1})$,
the variable $m_{eff}$ is defined as follows.

$$k_{j} = 1/\left(1+n\xi^{-1}s_{j}^{2}\eta_{j}^{-1} \right),
\quad j=1,2,...,p, \\ m_{eff} = \sum_{j=1}^{p}{\left(1-k_{j} \right)}.$$

$s_{j},\ j=1,2,...,p$ are the diagonal components of $X^{T}X$. For the zero 
components of $\beta$, $k_{j}$ is derived close to 1, and nonzero's $k_{j}$ is 
derived close to 0, so the variable $m_{eff}$ is called the effective number of 
nonzero coefficients. In this algorithm, $\delta$ is updated to set 
$s_{\delta} = ceiling(m_{eff})$. Adaptive probability is defined to satisfy 
Theorem 5(diminishing adaptation condition) of Roberts and Rosenthal (2007). 
at $T$th iteration,

$$p(T) = exp[p_{0} + p_{1}T],\quad p_{1} < 0,
\quad u \sim U(0, 1), \\ if\ u < p(T),\ accept\ \delta^{New},\ 
s_{\delta^{New}} = ceiling(m_{eff}).$$

The default is $p_{0} = 0$, $p_{1} = -4.6 \times 10^{-4}$, and 
under this condition, $p(10000) < 0.01$ is satisfied.

```{r fig.width = 6, fig.height= 4}
# approx_horseshoe with fixed default threshold.
approx_horseshoe_result <- approx_horseshoe(y, X, burn = 0, iter = 500, 
                                            auto.threshold = FALSE)

# modified approx_horseshoe with adaptive probability algorithm.
mapprox_horseshoe_result <- approx_horseshoe(y, X, burn = 0, iter = 500)

df2 <- data.frame(index = 1:100,
                  approx_BetaHat = approx_horseshoe_result$BetaHat[1:100],
                  approx_LeftCI = approx_horseshoe_result$LeftCI[1:100],
                  approx_RightCI = approx_horseshoe_result$RightCI[1:100],
                  mapprox_BetaHat = mapprox_horseshoe_result$BetaHat[1:100],
                  mapprox_LeftCI = mapprox_horseshoe_result$LeftCI[1:100],
                  mapprox_RightCI = mapprox_horseshoe_result$RightCI[1:100],
                  true_beta = true_beta[1:100])

# Estimation results of the approx_horseshoe.
ggplot(data = df2, aes(x = index, y = true_beta)) + 
  geom_point(size = 2) + 
  geom_point(aes(x = index, y = approx_BetaHat), size = 2, col = "red") +
  geom_errorbar(aes(ymin = approx_LeftCI, 
                    ymax = approx_RightCI), width = .1, col = "red") +
  labs(title = "95% Credible intervals of the approx_horseshoe", y = "beta")

# Estimation results of the mapprox_horseshoe.
ggplot(data = df2, aes(x = index, y = true_beta)) + 
  geom_point(size = 2) + 
  geom_point(aes(x = index, y = mapprox_BetaHat), 
             size = 2, col = "red") +
  geom_errorbar(aes(ymin = mapprox_LeftCI, 
                    ymax = mapprox_RightCI), 
                width = .1, col = "red") +
  labs(title = "95% Credible intervals of the modified_approx_horseshoe", 
       y = "beta")
```

In the case of exact_horseshoe, since the $D$ is used as is, it can always be 
thought of as $length(S) = 500$. On the other hand, since the method of setting 
the threshold is different in approx_horseshoe and modified approx_horseshoe, a 
difference occurs in $length(S) = s_{\delta}$, which causes a difference in the 
calculation speed of the algorithm.

```{r fig.width = 6, fig.height= 4}
exact_activeset <- rep(p, 500)
approx_activeset <- apply(approx_horseshoe_result$ActiveSet, MARGIN = 1, sum)
mapprox_activeset <- apply(mapprox_horseshoe_result$ActiveSet, MARGIN = 1, sum)

# active set plot
ggplot(data = data.frame(X = 1:500,
                         exact_activeset = exact_activeset,
                         approx_activeset = approx_activeset,
                         mapprox_activeset = mapprox_activeset)) +
  geom_line(mapping = aes(x = X, y = exact_activeset,
                          color = "exact")) +
  geom_line(mapping = aes(x = X, y = approx_activeset,
                          color = "approx"),
            alpha = 0.5) +
  geom_line(mapping = aes(x = X, y = mapprox_activeset,
                          color = "modified_approx"),
            alpha = 0.5) +
  scale_color_manual(name = "algorithm",
                     values = c("black", "red", "blue"), 
                     breaks = c("exact", "approx", "modified_approx"), 
                     labels = c("exact", "approx", "modified_approx"))
```

# References

Bhattacharya, A., Chakraborty, A., & Mallick, B. K. (2016). Fast sampling with 
Gaussian scale mixture priors in high-dimensional regression. Biometrika, 
asw042.

Johndrow, J., Orenstein, P., & Bhattacharya, A. (2020). Scalable Approximate 
MCMC Algorithms for the Horseshoe Prior. In Journal of Machine Learning 
Research (Vol. 21).

Piironen, J., & Vehtari, A. (2017). Sparsity information and regularization 
in the horseshoe and other shrinkage priors. Electronic Journal of 
Statistics, 11, 5018-5051.

Roberts G, Rosenthal J. Coupling and ergodicity of adaptive Markov chain Monte 
Carlo algorithms. J Appl Prob. 2007;44:458–475.
